AHABus \& FCORE:\\Modular Bus for High-Altitude Balloon Missions

### 2017-04-09 chapters/03-methodology.tex
Add GNSS task

FreeRTOS is a real-time kernel that provides a prioritised task model: each task
is given a function (usually, an event loop) to run and a priority level. At
regular intervals, the kernel establishes a list of all the tasks that must run,
and gives processor time to the one with the highest priority. Such a
deterministic system is desirable for applications like \acrshort{ahabus}, which
must run on embedded devices with limited computing power, and where certain
actions must imperatively run at regular intervals. \acrshort{fcore} was
designed around three main tasks, along with the standalone radio subsystem

\subsubsection{Payload Bus Task}
Each payload is assigned its own task, with a running
interval depending on how the payload was configured before flight. Each task
fetches data from its payload, packetises it and adds it to the radio queue.

@@

The \acrshort{gnss} task was designed to run at fixed intervals to parse the
\acrshort{gnss} receiver's output, and update the platform's cached
co-ordinates and altitude. This cached location is the one included in radio
packets.

>>>>

FreeRTOS is a real-time kernel that provides a prioritised task model: each task
is given a function (usually, an event loop) to run and a priority level. At
regular intervals, the kernel establishes a list of all the tasks that must run,
and gives processor time to the one with the highest priority. Such a
deterministic system is desirable for applications like \acrshort{ahabus}, which
must run on embedded devices with limited computing power, and where certain
actions must imperatively run at regular intervals. \acrshort{fcore} was
designed around three main tasks, along with the standalone radio subsystem.
Figure \ref{fig:fcore-arch} shows an overview of the the \acrshort{fcore}
architecture, and the component presented in the following sections.

\begin{figure}[H]
\includegraphics[width=0.6\textwidth]{fcore-arch.pdf}
\centering
\caption{\acrshort{fcore} architecture}
\label{fig:fcore-arch}
\end{figure}

\subsubsection{Payload Bus Task}

The payload bus subsystem was designed so that each payload connected to the
platform is given a unique identifier, which is also its address on the
\acrshort{i2c} bus. To avoid errors emanating from one payload to propagate
and take the whole payload bus offline, each payload is assigned a Payload
Bus Task.

A Bus Tasks runs at a fixed interval, configured at compile time depending on
the payload's priority level, as discussed in more details in section
\ref{sssec:conf-file}. The task establishes a connection with the payload using
the \acrshort{ahabus} payload bus protocol. If data is available, it is
collected into a temporary buffer and a packet with the platform's cached
geographical coordinates is created and sent to the radio submodule.

While the ESP8266 micro-controller does not have any hardware \acrshort{i2c}
support, the FreeRTOS SDK provides an implementation based on hardware timers
that was used for bus operations. As discussed in the results (section
\ref{ssec:results-bus}), the library had to be modified to extend the
\acrshort{i2c} clock stretch timeout to \SI{100000}{\micro\second}, so that
payloads with relatively slow processors had enough time to reply to commands.

Because all the payload bus tasks run at the same priority level, there was a
possibility that a payload's task could take control of the kernel while another
was still communicating with its own payload. To avoid this, the portion of the
code that handles communications over the bus and packetisation was marked
as critical, preventing FreeRTOS from triggering a context switch until all data
is collected.

@@

The \acrshort{gnss} task was designed to run at fixed intervals of
\SI{20}{\second} to parse the \acrshort{gnss} receiver's output, and update the
platform's cached co-ordinates and altitude. This cached location is the one
included in radio packets.

\acrshort{fcore} was designed to use \acrshort{gnss} receivers that comply with
the NMEA 0183 standard (\cite{Bekte2001}). In order to avoid the added risk of
writing a custom parser for an existing standard, the minmea library
(\cite{Moczek2016}) was chosen for its simplicity and reasonable size.

The NMEA sentences used to keep track of a receiver positions are called GGA.
Since receivers transmit multiple types of sentences every second, the
\acrshort{gnss} task was designed to attempt to parse up to five sentences
every time it runs, until a GGA sentence is found. If no GGA sentence is 
received, the task sets the \acrshort{gnss} health status to recovering to
indicate that transmitted positions might be out of date.
